### 2022-04-06 outputs/paper/paper.Rmd
update to paper

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(scales)
library(broom)

data <- readr::read_csv("../data/cleaned_data.csv")
data <- data %>%
  filter(demographic_info != "Don’t know/missing", survey_answer != "don't know")
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE,fig.width=8, fig.height=4}
age_group_diverging <- data %>%
  filter(demographic_type == "age group") %>%
  mutate(chances_of_getting_aids = if_else(survey_answer %in% c("no risk at all", "small"), -chances_of_getting_aids, chances_of_getting_aids)) %>%
  mutate(percent_label = abs(chances_of_getting_aids)) %>%
  mutate(percent_label = percent(chances_of_getting_aids / 100, accuracy = 1))

age_group_diverging %>%
  filter(gender == "male") %>%
  ggplot(aes(
    x = demographic_info,
    y = chances_of_getting_aids,
    fill = survey_answer
  )) +
  geom_col() +
  geom_text(aes(label = percent_label),
    position = position_stack(vjust = 0.5),
    color = "white",
    fontface = "bold"
  ) +
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(
    breaks = c("no risk at all", "small", "moderate", "great"),
    values = c(
      "no risk at all" = "darkslategray4",
      "small" = "darkslategray2",
      "moderate" = "deepskyblue2",
      "great" = "deepskyblue4"
    )
  ) +
  labs(
    title = "What is your chances of getting AIDS?",
    x = "age group",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    panel.grid = element_blank(),
    legend.position = "top"
  )
age_group_diverging %>%
  filter(gender == "female") %>%
  ggplot(aes(
    x = demographic_info,
    y = chances_of_getting_aids,
    fill = survey_answer
  )) +
  geom_col() +
  geom_text(aes(label = percent_label),
    position = position_stack(vjust = 0.5),
    color = "white",
    fontface = "bold"
  ) +
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(
    breaks = c("no risk at all", "small", "moderate", "great"),
    values = c(
      "no risk at all" = "pink4",
      "small" = "pink2",
      "moderate" = "brown2",
      "great" = "brown4"
    )
  ) +
  labs(
    title = "What is your chances of getting AIDS?",
    x = "age group",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    panel.grid = element_blank(),
    legend.position = "top"
  )
age_group_distribution <- age_group_diverging %>% distinct(population_count, .keep_all = TRUE)
age_group_distribution %>% 
  ggplot(aes(fill = gender, y = population_count, x = demographic_info)) +
  geom_bar(position = "dodge", stat = "identity") +
  coord_flip() +
  scale_x_discrete() +
  scale_fill_manual(
    values = c(
      "male" = "darkslategray3",
      "female" = "pink3"
    )
  ) +
  labs(
    title = "Survey Respondent Distribution",
    x = "age group",
    y = "survey respondents",
    fill = NULL
  ) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    panel.grid = element_blank(),
    legend.position = "top"
  )

```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

```{r f, echo = FALSE, warning=FALSE}
data1 <- data %>% filter( demographic_type == "age group")

scatterPlot1 <- ggplot(data1, aes(demographic_info, chances_of_getting_aids, color = survey_answer)) +
  geom_point() + geom_jitter()+ facet_wrap(~gender)+
  stat_smooth(aes(group = survey_answer), method = "loess", se = FALSE, lty='dashed') +
  labs(
    title = "Survey Respondent Distribution",
    x = "age group",
    y = "survey respondents",
    fill = NULL
  )+scale_color_brewer(
    palette="Spectral")+
  theme_minimal()
scatterPlot1
```
# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References